Measuring contrast processing in the visual system using SSVEP (and VEP?)

### 2024-03-20 review.qmd
Writing

>>>>

In early work on infants, the ‘sweep VEP’ paradigm was developed to measure SSVEP activity in response to a stimulus that changes its properties throughout a trial, for example gradually increasing in contrast or spatial frequency. Other studies keep the stimulus properties constant for the entire trial, which avoids serial effects and makes the analysis more straightforward.

Because it provides a direct read-out of neural population activity, the SSVEP signal can reveal key features of neural signal transduction. By parametrically varying the peak stimulus contrast, a monotonically increasing contrast response function can be measured. This corresponds closely to similar functions measured in single unit studies, or those measuring local field potentials invasively in the cortex. However SSVEP has the advantage that it is non-invasive, and so can be measured in awake, behaving human participants. An early use of the contrast response function was to provide an automated estimate of contrast sensitivity, without requiring behavioural responses from the participant. This was achieved by measuring the contrast response function, and extrapolating back along the function to estimate its intercept with the x-axis, which was shown to correspond with psychophysically measured detection thresholds (REFS). This general approach revealed much about the development of visual abilities in infants (Braddick), as well as deficits in neurological disease. However, psychophysical measurements remain the gold standard, and do not always correspond closely with SSVEP estimates in individuals with normal adult visual systems.
